---
title: "Project 2"
author: "Nancie Kung, Calvin Raab, David Collier and Eitan Shimonovitz"
date: "5/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tseries)
library(forecast)
library(dynlm)
library(lmtest)
```

# Introduction 
In this project, we will be analyzing Hass Avocados. Our objective is to use a time series to explain how the average price and total volume of Hass avocados have changed over time. First, we explored the data by looking at measures such as the changes in price and volume and the popularity of avocados by region. We then created an AR and ARDL model to examine the patterns in price and volume in the past. We also used these models to make predictions about how price and volume would change in the future. Below, we show our analysis. 

# Description of the Data
Our project uses historical data on avocado prices and sales volume in U.S. markets between the years 2015 and 2018. This data is based on the weekly retail sales of Hass avocados reported by retailers' cash registers. It contains an aggregation of data from multiple locations across the United States and multiple types of retail outlets. The average price is the per unit cost for each avocado and the Product Lookup code indicates the total number sold for a given type of Hass avocado. The following variables are included in the data set:

• Date - the date of the observation  
• AveragePrice - the average price of a single avocado  
• Type - conventional or organic  
• Year - the year  
• Region - the city or region of the observation  
• Total Volume - total number of avocados sold  
• Hass.Small - total number of avocados with PLU 4046 sold (Small Hass Avocados)  
• Hass.Large - total number of avocados with PLU 4225 sold (Large Hass Avocados)  
• Hass.Extra.Large - total number of avocados with PLU 4770 sold (Extra Large Hass Avocados)  

# Load Data
```{r}
avocados <- read.csv("avocado.csv")
attach(avocados)
avocados <- avocados %>%
  rename(
    Hass.Small = X4046,
    Hass.Large = X4225,
    Hass.Extra.Large = X4770
  )
```

### Creating A Time Series 
```{r}
library(dplyr)

avocados <- avocados %>% arrange(region, type, Date)

avocados_us_conventional <- avocados %>% dplyr::filter(region == "TotalUS", type == "conventional")

avocados_us_conv.ts <- ts(avocados_us_conventional, frequency = 52, start = c(2015,1), end = c(2018,13))

avocados_us_organic <- avocados %>% dplyr::filter(region == "TotalUS", type == "organic")
avocados_us_org.ts <- ts(avocados_us_organic, frequency = 52, start = c(2015,1), end = c(2018,13))
```

# Exploratory Analysis 
When creating our time series above we chose to us "TotalUS" as the region and "conventional" as the type for we believed that was the best representation of our data. To back up our decision we looked at a breakdown of this dataset to see what percent was conventional versus what percent was organic. Here it can be seen that 97% of the avocados recorded in this dataset are conventional, hence why we chose to use it. 
```{r}
volume_check <- avocados %>% 
  group_by(type) %>% 
  summarise(avgerage_volume=mean(Total.Volume))  %>% 
  mutate(Percent=proportions(avgerage_volume) * 100) 

volume_check
```


### Average Price of Avocados Over Time
```{r}
plot(avocados_us_conv.ts[,3], ylab = "Avocado Price", xlab = "Date")
```


### Difference In Number of Avocados Sold Overtime for Different Size Haas Avocados 
```{r}
ts.plot(avocados_us_conv.ts[,5:7]/1000000, gpars=list(col=rainbow(4)),ylab = "Avocados Sold (Millions)", xlab = "Date", main = "Number of Avocados Sold")
legend(2015, 23, legend=c("S Avocados", "L Avocados", "XL Avocados"),
       col=c("red", "green", "cyan"), lty=1:2, cex=0.6)
```

### Difference In Number of Avocados Sold Over Time for Conventional vs Organic Avocados
```{r}
par(mfrow=c(2,2))
plot((avocados_us_conv.ts[,4] / 1000000), ylab = "Avocados Sold (Millions)", xlab = "Date", main = "Number of Conventional Avocados Sold", col = "brown")
plot((avocados_us_org.ts[,4] / 1000000), ylab = "Avocados Sold (Millions)", xlab = "Date", main = "Number of Organic Avocados Sold", col = "green")
```

Conventional avocados show strong seasonality, with spikes in the beginning of the year, while organic avocados show a consistent upwards trend from 2015 - 2018. Organic avocados appear to show their seasonality through severe dips at the end of the year, in the beginning of winter.

### Average Price By Region 
```{r}
avg_price_by_region <- aggregate(x = avocados$AveragePrice,
          by = list(avocados$region),
          FUN = mean)

#Top 10 Most Expensive Regions
slice(arrange(avg_price_by_region, desc(x)), 1:10)

#Top 10 Least Expensive Regions
slice(arrange(avg_price_by_region, x), 1:10)
```


# Data Analysis/Model

```{r}
y <- avocados_us_conv.ts[, "Total.Bags"]
ar_mod1 <- ar(y, aic = FALSE, order.max=2, method="ols")
summary(ar_mod1)
ar_mod1
forecast(ar_mod1, 52)
```

### AR Process 
In order to see if this avocado data is cyclical, we will look at the ACF of both the Average Price and the Total Volume of our avocados. We set the lag to be maxed at 52, because there are 52 weeks in a year. From the ACF is can be seen that in the middle of the year there isn't much correlation, however near the end of the year we start to see significant statistical correlation. This tells us that there appears to be a cyclical, yearly relationship between our data. This means that data from 12 months ago can help predict the data of today. 
```{r}
library(tseries)
library(forecast)
AveragePrice.ts <- avocados_us_conv.ts[,3]
TotalVolume.ts <- avocados_us_conv.ts[,4]
tsdisplay(AveragePrice.ts, lag.max = 52)
tsdisplay(TotalVolume.ts, lag.max = 52)
```
The pattern of a steadily decreasing ACF with a single spike at lag = 1 for the PACF shows the pattern of an AR(1) process for average price. In addition, the pattern of a steadily decreasing ACF with multiple spikes at lag = 1,2,3,13,14 suggest a higher order AR process for average volume. 

### Prediction with our AR model 
Here is an AR prediction model. We built a model to predict average price. The data we fed into our model was subsetted so we could use the final 5 results to test how accurate our model is. From the results below it can be seen that our prediction model did a good job and all of our confidence intervals created by the AR model contained the actual prices. 
```{r}
avo.ar <- ar(AveragePrice.ts[1:164], aic = FALSE, order.max=1, method = "ols")
summary(avo.ar)
print(avo.ar) # Here you can see the two coefficients of the two lags 

plot(AveragePrice.ts) 
lines(time(AveragePrice.ts)[1:164],AveragePrice.ts[1:164] - avo.ar$resid, col = "green")

plot(forecast(avo.ar, 5))

# Forecast 5 steps-ahead 
forecast(avo.ar, 5)

# Compare to the actual next 5
AveragePrice.ts[165:169]
```


### Using Seasonality With AR(1) Model

The ACF and PACF for average price show clear seasonality year-over-year, which is to be expected with produce such as avocados. As such, we investigated including seasonality in our model to account for this effect, with AR(1) and MA(1) seasonality.
```{r}
seasonal_model <- Arima(AveragePrice.ts, order = c(1,0,0), seasonal = list(order=c(1,0,1), period = 52))

summary(seasonal_model)

plot(AveragePrice.ts)
lines(fitted(seasonal_model), col = "green")

AveragePrice.ts
```

We can see that the fitted values follow the data very closely. Next, we use the seasonal model to predict average price for the next year.

```{r}
plot(forecast(seasonal_model,52))
```

We see that the predicted price follows the seasonal trend of rising thoughout the year, then falling at the end of the year as we expect. This matches up with our knowledge of avocados being cheapest in the winter months and rising in price during the rest of the year. We also looked into forecasting with the built-in exponential smoothing capabilities of the forecast() function.

```{r}
plot(forecast(AveragePrice.ts, 52))
```

The exponentially smoothed prediction also follows the seasonal trend, but predicts that the trend of average price will actually decrease in the next year, which we do not believe to be likely. 


### Buidling an ARDL model 
Below is an ARDL(4,4) model. We built this model to see the statistical significance lags of both average price and total volume would have on predicting average price. We chose a lag of 4, for that represents one months worth of lags. The findings here are interesting in that lags of average price appeared to be less statistically significant than lags of volume. It appears that average price from 1 week ago, current total volume, and total volume from 1 week ago are statistically significant in explaining the current average price.
```{r}
# ARDL(4,4)
avo.ardl <- dynlm(AveragePrice.ts ~ L(AveragePrice.ts,1:4) + L(TotalVolume.ts, 0:4))
summary(avo.ardl)
```

### Testing for Serially Correlation

Here we test whether the ARDL model above violates the assumption that the errors are serially correlated. Looking at the ACF and PACF plots for the model's residuals, we do not see any distinctive pattern which suggests that the errors are not serially correlated. The Breusch-Godfrey test for higher order serial correlation also suggests that there is no serial correlation in the model, since the high p-value means that we fail the reject the null. This means that we do not have to correct for serial correlation.

```{r}
library(tseries)
tsdisplay(avo.ardl$residuals)
bgtest(avo.ardl, order=1, type="F", fill=0)
```

### Testing Instrumental Variables 
Using our subject mater expertise on avocados we thought to test total bags as an instrumental variable. The reason for this is we believed total bags to be a good indicator of total volume, but not necessarily of price. We created that IV test below. From our new model that takes into consideration total bags, it can be seen that there isn't much change in the significance of our parameters, suggesting that total bags is a weaker IV. 
```{r}
# This will only run if the second 
# Before testing total bags as an IV 
no.iv.mod <- lm(AveragePrice ~ Total.Volume + Hass.Small + Hass.Large + Hass.Extra.Large, data = avocados_us_conventional)
summary(no.iv.mod)

# Total Bags is the IV we are testing 
library(AER)
total.bags.iv <- ivreg(AveragePrice ~ Total.Volume + Hass.Small + Hass.Large + Hass.Extra.Large | Hass.Small + Hass.Large + Hass.Extra.Large + Total.Bags, data = avocados_us_conventional)
summary(total.bags.iv)

```

# Conclusions
Based on our analysis, we make the following key conclusions:

• Both average price and total volume follow an AR process, in which present values of the time series are related to past values.  
• Current average price can be explained by recent lags in average price and total volume.  
• We can predict the the short-term future average price given that it follows a seasonal pattern.  
• The price of avocados throughout the year follows a consistent seasonal pattern of rising until a peak in early fall, then dropping to its lowest point at the end of the year. As such, we would advise consumers to avoid purchasing avocados in late summer and early fall, then get plenty of avocados during the winter for holiday and Super Bowl guacamole.   
• This seasonal effect was weak in 2015 but became more prominent in following years. It's possible that external weather factors during the later years reduced the yield of out-of-season avocados, though the true reason is unclear in our data.  

# Future Work
For our future work, we believe we can improve on our model by testing the performance of our model. We can do this by using cross validation, in which we would divide our data into training and testing sets in order to evaluate how well our model performed. This evaluation can inform us if we should restructure our model to improve its accuracy.

We also would like to further investigate the difference between conventional and organic avocados. While conventional makes up the vast majority of the marketplace, we saw that the trends were different for the two types, so it would be interesting to look into the effect of the increasing popularity of organic produce on the avocado market.

In addition, we believe that it would useful to account for other variables not included in this dataset that could explain changes in avocado price and volume. For example, we could integrate information on weather patterns and economic conditions to provide context for our interpretations. This would help us explain the trends in the data within a bigger picture. 

# References

Avocado data from kaggle: https://www.kaggle.com/neuromusic/avocado-prices
